Relative parabolic factorizations of Chevalley groups

### 2015-09-01 congrsubgr-boundgen.tex
Added one more auxiliary lemma to the introduction

\begin{lemma}\label{lemma:srRI1_width} In the assumptions of Lemma~\ref{lemma:srRI1} the width of $\E(\Phi,R,I)$ with respect to $\mathcal{Z}(\Pi)$ does not exceed $3\left|\Phi^+\right|+2\rk\Phi-1$. \end{lemma}
\begin{proof}
Take an element $g\in\E(\Phi,R,I)$ and decompose it into $g=u_1 h v_2 u_3$, where $h\in\Hh(\Phi,R,I)$, $u_1,u_3\in\U(\Phi,I)$, $v_2\in\U(\Phi^-,I)$. 
Write $h=\prod_{i=1}^\ell h_{\alpha_i}(\varepsilon_i)$, $\varepsilon\in1+I$. 
Each $h_{\alpha_i}(\varepsilon_i)$ decomposes into $h_{\alpha_i}(\varepsilon_i) = x_{\alpha_i}(*) z_{-\alpha_i}(*,*) x_{-\alpha_i}(*)$ 
 (see Lemma~\ref{lemma:rel-tor-elementary}), and since the torus normalizes each of $X_\alpha(I)$ (see formula~\ref{rel:R6}), we have a decomposition
\[ g\in\U(\Phi,I)\cdot\prod_{i=1}^\ell\bigl(x_{\alpha_i}(*)z_{-\alpha_i}(*,*)\bigr)\cdot \U^-(\Phi,I) \U(\Phi,I), \]
and the estimate follows. \end{proof}

>>>>

\begin{lemma}\label{lemma:srRI1_width} In the assumptions of Lemma~\ref{lemma:srRI1} the width of $\E(\Phi,R,I)$ with respect to $\mathcal{Z}(\Pi)$ does not exceed $3\left|\Phi^+\right|+2\rk\Phi-1$. \end{lemma}
\begin{proof}
Take an element $g\in\E(\Phi,R,I)$ and decompose it into $g=u_1 h v_2 u_3$, where $h\in\Hh(\Phi,R,I)$, $u_1,u_3\in\U(\Phi,I)$, $v_2\in\U(\Phi^-,I)$. 
Write $h=\prod_{i=1}^\ell h_{\alpha_i}(\varepsilon_i)$, $\varepsilon\in1+I$. 
Each $h_{\alpha_i}(\varepsilon_i)$ decomposes into $h_{\alpha_i}(\varepsilon_i) = x_{\alpha_i}(*) z_{-\alpha_i}(*,*) x_{-\alpha_i}(*)$ 
 (see Lemma~\ref{lemma:rel-tor-elementary}), and since the torus normalizes each of $X_\alpha(I)$ (see formula~\ref{rel:h-w}), we have a decomposition
\[ g\in\U(\Phi,I)\cdot\prod_{i=1}^\ell\bigl(x_{\alpha_i}(*)z_{-\alpha_i}(*,*)\bigr)\cdot \U^-(\Phi,I) \U(\Phi,I), \]
and the estimate follows. \end{proof}

### 2015-09-01 paper.tex
Added one more auxiliary lemma to the introduction

Now we are ready to state a recent theorem of A.~Stepanov (see~\cite[Theorem~3.4]{S}).  
\begin{thm*}\label{thm:Stepanov}
Let $\Phi$ be a reduced irreducible root system of rank $\geq 2$ and let $S \subset \Phi$ be any parabolic subset of roots.
Then $\E(\Phi, R, I)$ is generated as an abstract group by $\mathcal{Z}(\Sigma_S)$.
\end{thm*}

@@

Let $J\subseteq \Pi$ be a subset of simple roots. 
For a root $\beta = \sum d_i\alpha_i \in \Phi$ set $\shape(J, \beta):=\sum\limits_{i\in J}d_i \alpha_i$.
Set $$\Delta_J = \{\alpha \in \Phi \mid \shape(J, \alpha)=0\},\quad \Sigma^\pm_J = \{\alpha \in \Phi \mid \shape(J, \alpha)\in\Phi^\pm \}.$$
Clearly, $\Delta_J$ is a reductive subset of roots while $\Sigma^\pm_J$ are special subsets.
For two disjoint subsets $I, J\subseteq \Pi$ one has $\Sigma^\pm_{I \cup J} = \Sigma^\pm_I\cup\Sigma^\pm_J$ while $\Delta_{I\cup J} = \Delta_I \cap \Delta_J$.

@@

\subsection{Elementary subgroup}\label{sec:elementary}
Recall from~\cite{St78, VP} that for $\alpha\in \Phi$, $\xi\in R$ one can define certain elements $x_{\alpha}(\xi)$ of $\G_{sc}(\Phi, R)$ called {\it elementary root unipotents}.
These elements satisfy the well-known Steinberg relations:
\begin{align}
& \phantom{\left[{}\vphantom{x_\beta(t)}\right.}
x_\alpha(s) x_\alpha(t) = x_\alpha(s+t), \label{rel:add}\\
& \left[x_\alpha(s),  x_\beta(t)\right] = \prod x_{i\alpha + j\beta}\left(N_{\alpha\beta ij}\, s^i t^j\right),\quad \alpha\neq-\beta, \quad N_{\alpha\beta ij}\in\mathbb{Z}. \label{rel:CCF}
\end{align} 
The indices $i$, $j$ in the above formula range over all positive natural numbers such that $i\alpha + j\beta\in\Phi$.
The integers $N_{\alpha\beta ij}$ are called {\it structure constants} of the Chevalley group $\G(\Phi,R)$ and depend only on $\Phi$.

For $\varepsilon\in R^*$ set $w_\alpha(\varepsilon) = x_\alpha(\varepsilon) x_{-\alpha}(-\varepsilon^{-1}) x_{\alpha}(\varepsilon).$
If $\rk(\Phi)\geqslant 2$ the following relation is a consequence of \ref{rel:add}--\ref{rel:CCF}:
\begin{equation}\label{rel:R3}
w_\alpha(\varepsilon) x_{\beta}(\xi) w_\alpha(\varepsilon)^{-1} =
x_{\sigma\ssub{\alpha}\beta} \left(\eta_{\alpha, \beta}\cdot\varepsilon^{-\langle\beta,\alpha \rangle}\xi\right),\quad \varepsilon\in R^*,\ \xi\in R.
\end{equation}
where coefficients $\eta_{\alpha, \beta} = \pm 1$ (see~\cite[\S13]{VP} for more details on $\eta_{\alpha, \beta}$).

Define semisimple root elements $h_\alpha(\varepsilon)$, $\alpha\in\Phi$, $\varepsilon\in R^*$ as $h_\alpha(\varepsilon)=w_\alpha(\varepsilon)w_\alpha(-1)$. Then
\begin{equation}\label{rel:R6}
h_\alpha(\varepsilon)x_\beta(\xi)h_\alpha(\varepsilon)^{-1} = x_\beta\left(\varepsilon^{\langle\beta,\alpha\rangle}\xi\right),\quad \varepsilon\in R^*,\ \xi\in R.
\end{equation}

@@

In case $\Phi=\rC_\ell$ the long and short root unipotenrs are
\begin{align*}
& x_{i,-i}(\xi)=e+\xi e_{i,-i}, \\ 
& x_{i,j}(\xi)=e+\xi e_{ij}-\varepsilon_i\varepsilon_j\xi e_{-j,-i},\quad i\neq\pm j.
\end{align*}

@@

\subsection{$\K_1$-functor modeled on Chevalley groups}

Recall that by Taddei's theorem $\E(\Phi, R)$ is a normal subgroup of $\G_{sc}(\Phi, R)$ provided $\Phi$ is an irreducible root system of rank $\geqslant 2$ (see~\cite{Ta}).

>>>>

Now we are ready to state a recent result of A.~Stepanov.  
\begin{prop}[{\cite[Theorem~3.4]{S}}] \label{prop:Stepanov}
Let $\Phi$ be a reduced irreducible root system of rank $\geq 2$ and let $S \subset \Phi$ be any parabolic subset of roots.
Then $\E(\Phi, R, I)$ is generated as an abstract group by $\mathcal{Z}(\Sigma_S)$.
\end{prop}

@@

Let $J\subseteq \Pi$ be a subset of simple roots. 
For a root $\beta = \sum m_i(\beta)\alpha_i \in \Phi$ set $\shape(J, \beta):=\sum\limits_{i\in J} m_i(\beta) \alpha_i$.
Set $$\Delta_J = \{\alpha \in \Phi \mid \shape(J, \alpha)=0\},\quad \Sigma^\pm_J = \{\alpha \in \Phi \mid \shape(J, \alpha)\in\Phi^\pm \}.$$
Clearly, $\Delta_J$ is a reductive subset of roots while $\Sigma^\pm_J$ are special subsets.
For two disjoint subsets $I, J\subseteq \Pi$ one has $\Sigma^\pm_{I \cup J} = \Sigma^\pm_I\cup\Sigma^\pm_J$ while $\Delta_{I\cup J} = \Delta_I \cap \Delta_J$.

@@

\subsection{Elementary subgroup}\label{sec:elementary}
Recall from~\cite{St78, VP} that for $\alpha\in \Phi$, $\xi\in R$ one can define certain elements $x_{\alpha}(\xi)$ of $\G_{sc}(\Phi, R)$ called {\it elementary root unipotents}.
These elements satisfy the well-known Steinberg relations:
\begin{align}
& \phantom{\left[{}\vphantom{x_\beta(t)}\right.}
x_\alpha(s) x_\alpha(t) = x_\alpha(s+t), \label{rel:add}\\
& \left[x_\alpha(s),  x_\beta(t)\right] = \prod x_{i\alpha + j\beta}\left(N_{\alpha\beta ij}\, s^i t^j\right),\quad \alpha\neq-\beta, \quad N_{\alpha\beta ij}\in\mathbb{Z}. \label{rel:CCF}
\end{align} 
The indices $i$, $j$ in the above formula range over all positive natural numbers such that $i\alpha + j\beta\in\Phi$.
The integers $N_{\alpha\beta ij}$ are called {\it structure constants} of the Chevalley group $\G(\Phi,R)$ and depend only on $\Phi$.
We refer the reader to~\cite[\S14]{VP} for basic properties of structure constants.

For $\varepsilon\in R^*$ set $w_\alpha(\varepsilon) = x_\alpha(\varepsilon) x_{-\alpha}(-\varepsilon^{-1}) x_{\alpha}(\varepsilon).$
If $\rk(\Phi)\geqslant 2$ the following relation is a consequence of \ref{rel:add}--\ref{rel:CCF}:
\begin{equation}\label{rel:R3}
w_\alpha(\varepsilon) x_{\beta}(\xi) w_\alpha(\varepsilon)^{-1} =
x_{\sigma\ssub{\alpha}\beta} \left(\eta_{\alpha, \beta}\cdot\varepsilon^{-\langle\beta,\alpha \rangle}\xi\right),\quad \varepsilon\in R^*,\ \xi\in R.
\end{equation}
where $\eta_{\alpha, \beta}=\pm 1$. The coefficients $\eta_{\alpha, \beta}$ can be expressed through structure constants (cf.~\cite[\S13]{VP}).
For a subset $\Psi \subseteq \Phi$ denote by $\widetilde{W}(\Psi)$ the subgroup of $\E(\Phi, R)$ generated by all $w_{\alpha}(1)$, $\alpha \in \Psi$.
\begin{lemma} \begin{thmlist} \item For every roots $\alpha, \beta \in \Phi$ having the same length there exists $w \in \widetilde{W}(\Phi)$ such that $X_{\alpha}(I)^w = X_\beta(I)$.
 \item The elementary group $\E(\Phi, R)$ (resp. $\EP_r(R)$, $1\leq r\leq \ell$) is generated as an abstract group by root subgroups $X_{\alpha}(R)$, $\alpha \in S$,
       where $S = \Pi \cup -\Pi$ (resp. $S = \Pi \cup -\Pi \setminus \{ -\alpha_r \}$).
 \item Assume that for a simple root $\alpha_s\in \Pi$ one has $m_s(\alpha_\mathrm{max}) = 1$.
       Then for $\alpha, \beta \in \Sigma^\pm_s$ having the same length there exists $w\in \widetilde{W}(\Delta_s)$ such that $X_\alpha(I)^w = X_\beta(I)$. \end{thmlist} \end{lemma}
\begin{proof}
 The first statement of the lemma follows from identity~\ref{rel:R3} and the well-known fact that the Weyl group $W(\Phi)$ acts transitively on the set of roots of the same length.

 The prove the second statement observe that the subgroup $\langle X_\alpha(R) \mid \alpha \in S \rangle$ contains $\widetilde{W}(\Phi)$ (resp. $\widetilde{W}(\Delta_s)$).
 %TODO: FINISH PROOF
\end{proof}

Define semisimple root elements $h_\alpha(\varepsilon)$, $\alpha\in\Phi$, $\varepsilon\in R^*$ as $h_\alpha(\varepsilon)=w_\alpha(\varepsilon)w_\alpha(-1)$.
The elements $h_\alpha(\varepsilon)$ satisfy the following relations:
\begin{align}
 [ h_\alpha(\varepsilon_1), h_\beta(\varepsilon_2)] & = 1, & \alpha, \beta \in \Phi,\ \varepsilon_1, \varepsilon_2 \in R^*  \label{rel:h-comm} \\
 h_\alpha(\varepsilon_1 \varepsilon_2) & = h_\alpha(\varepsilon_1) h_\alpha(\varepsilon_2) & \alpha \in \Phi,\ \varepsilon_1, \varepsilon_2 \in R^* \label{rel:h-mult} \\
 h_\alpha(\varepsilon)x_\beta(\xi)h_\alpha(\varepsilon)^{-1} & = x_\beta\left(\varepsilon^{\langle\beta,\alpha\rangle}\xi\right) & \alpha, \beta \in \Phi,\ \varepsilon \in R^*, \xi \in R \label{rel:h-w}
\end{align}

@@

In case $\Phi=\rC_\ell$ the long and short root unipotents are
\begin{align*}
& x_{i,-i}(\xi)=e+\xi e_{i,-i}, \\ 
& x_{i,j}(\xi)=e+\xi e_{ij}-\varepsilon_i\varepsilon_j\xi e_{-j,-i},\quad i\neq\pm j.
\end{align*}

@@

\subsection{$\K_1$-functor modeled on Chevalley groups}
Recall that by Taddei's theorem $\E(\Phi, R)$ is a normal subgroup of $\G_{sc}(\Phi, R)$ provided $\Phi$ is an irreducible root system of rank $\geqslant 2$ (see~\cite{Ta}).
